---
title: "Beef Modeling"
author: "Mykola Dereva"
date: "8/2/2021"
output: html_document
---



```{r}
rm(list = ls())
```

```{r}
d <- readRDS("data/clean_data.rds")
```


# Modeling 

```{r message=FALSE}
library(apt)
library(urca)
library(dplyr)
```
There is a conflict with the dplyr lag() so I have to change it to the default 
function, so the apt package works properly 

```{r}
lag <- stats::lag
```



```{r}
product <- "Beef"

cons <- x <- d %>% 
  filter(Type == "Consumer",
         Product == product) %>% 
  select(Price) %>% 
  log() %>% 
  ts(start = c(2013, 1), frequency = 12)
  
prod <- y <- d %>% 
  filter(Type == "Producer",
         Product == product) %>% 
  select(Price) %>% 
  log() %>% 
  ts(start = c(2013, 1), frequency = 12)

x.name <- "Consumer"
y.name <- "Producer"
```

```{r}
par(mfrow=c(2,1))
plot(x)
plot(y)
```

### cointegration Tests

```{r}
ca.jo(x = cbind(x,y),
      type = "eigen",
      ecdet = "none",
      K = 2)
```




### EG contegration


```{r}
LR <- lm(y ~ x)
summary(LR)
```


residuals
```{r}
ry <- ts(residuals(LR), start=start(x), end=end(x), frequency = 12)
plot(ry)
```


```{r}
eg <- ur.df(ry, type = c("none"), lags=1)
summary(eg)
```

we are primarily interested in the residuals from this equation, which would suggest that the variables are cointegrated if the residuals are stationary.



```{r}
plot(eg)
```

Autocorrelation of residuals 

```{r}
eg4 <- Box.test(eg@res, lag = 4, type = "Ljung")
eg8 <- Box.test(eg@res, lag = 8, type="Ljung")
eg12 <- Box.test(eg@res, lag = 12, type="Ljung")

```

```{r}
eg4; eg8; eg12
```


# TAR + Cointegration


### TAR

```{r}
t3 <- ciTarThd(y=y, x=x, model="tar", lag=0)
```

```{r}
(th.tar <- t3$basic)
```

```{r}
plot(t3)
```



Choose threshold

```{r}
for (i in 1:12) { # 20 seconds
  t3a <- ciTarThd(y=y, x=x, model="tar", lag=i)
  th.tar[i+2] <- t3a$basic[,2]
}
```



```{r}
th.tar
```


### MTAR treshold

```{r}
t4 <- ciTarThd(y=y, x=x, model="mtar", lag=0)
(th.mtar <- t4$basic)
```



```{r}
plot(t4)
```


```{r}
for (i in 1:12) {
  t4a <- ciTarThd(y=y, x=x, model="mtar", lag=i)
  th.mtar[i+2] <- t4a$basic[,2]
}
th.mtar
```



The following threshold values are specific to this data. They HAVE TO be
revised for another data set. Otherwise, various errors will occur.

```{r}
(t.tar <- th.tar$V3[2])    #Best threshold for tar
(t.mtar <- th.mtar$V3[2])
```


## Lag selection 

```{r}
# max lag 
mx <- 12
```



### tar, thresold 0
```{r}
g1 <- ciTarLag(y=y, x=x, model="tar", maxlag=mx, thresh = 0)
g1
```
#### MTAR, threshlod 0
```{r}
(g2 <-ciTarLag(y=y, x=x, model="mtar",maxlag=mx, thresh=0))
```


```{r}
plot(g1)
```


```{r}
plot(g2)
```




```{r}
(g3 <-ciTarLag(y=y, x=x, model="tar", maxlag=mx, thresh=t.tar))

```


```{r}
(g4 <-ciTarLag(y=y, x=x, model="mtar",maxlag=mx, thresh=t.mtar))

```


# Table 3 Results of EG and threshold cointegration tests

Choose Lag
```{r}
vv <- 0
```

```{r}
(f1 <- ciTarFit(y=y, x=x, model="tar", lag=vv, thresh=0 ))
```


```{r}
(f2 <- ciTarFit(y=y, x=x, model="tar", lag=vv, thresh=t.tar ))
```



```{r}
(f3 <- ciTarFit(y=y, x=x, model="mtar", lag=vv, thresh=0 ))
```


```{r}
(f4 <- ciTarFit(y=y, x=x, model="mtar", lag=vv, thresh=t.mtar))
```


#### summary table

```{r}
r0 <- cbind(summary(f1)$dia, summary(f2)$dia,
            summary(f3)$dia, summary(f4)$dia)

diag <- r0[c(1:4, 6:7, 12:14, 8, 10, 9, 11), c(1,2,4,6,8)]

rownames(diag) <- 1:nrow(diag)
diag
```



```{r}
e1 <- summary(f1)$out
e2 <- summary(f2)$out
e3 <- summary(f3)$out
e4 <- summary(f4)$out

```


```{r}
ee <- list(e1, e2, e3, e4)
vect <- NULL


for (i in 1:4) {
  ef <- data.frame(ee[i])
  vect2 <- c(paste(ef[3, "estimate"], ef[3, "sign"], sep=""),
  paste("(", ef[3, "t.value"], ")", sep=""),
  paste(ef[4, "estimate"], ef[4, "sign"], sep=""),
  paste("(", ef[4, "t.value"], ")", sep=""))
  vect <- cbind(vect, vect2)
}

```


```{r}
item <- c("pos.coeff","pos.t.value", "neg.coeff","neg.t.value")
ve <- data.frame(cbind(item, vect))
colnames(ve) <- colnames(diag)
( res.CI <- rbind(diag, ve)[c(1:2, 14:17, 3:13), ] )
rownames(res.CI) <- 1:nrow(res.CI)

res.CI
```


```{r}
library(xtable)
```

### save summary table
```{r}
table <- xtable(res.CI)
print(table, type = "html",
      file = paste("tables/", product, ".doc", sep = ""))
```





# APT + ECM

```{r}
(sem <- ecmSymFit(y=prod, x=cons, lag=4))  # why lag 4?
```





```{r}
aem <- ecmAsyFit(y=prod, x=cons, lag=4, model="mtar", split=TRUE, thresh=t.mtar)
aem
```


```{r}
ccc <- summary(aem)

coe <- cbind(as.character(ccc[1:19, 2]),
    paste(ccc[1:19, "estimate"], ccc$signif[1:19], sep=""), ccc[1:19, "t.value"],
    paste(ccc[20:38,"estimate"], ccc$signif[20:38],sep=""), ccc[20:38,"t.value"])

colnames(coe) <- c("item",
                   paste(x.name, ".est", sep = ""),
                   paste(x.name, ".t", sep = ""),
                   paste(y.name, ".est", sep = ""),
                   paste(y.name, ".t", sep = "")
                   )

coe
```


```{r}
edia <- ecmDiag(aem, 3)
ed <- edia[c(1,6,7,8,9), ]
ed2 <- cbind(ed[,1:2], "_", ed[,3], "_")
colnames(ed2) <- colnames(coe)
ed2
```


```{r}
tes <- ecmAsyTest(aem)$out
tes2 <- tes[c(2,3,5,11,12,13,1), -1]

tes3 <- cbind(as.character(tes2[,1]),
    paste(tes2[,2], tes2[,6], sep=''), paste("[", round(tes2[,4],2), "]", sep=''),
    paste(tes2[,3], tes2[,7], sep=''), paste("[", round(tes2[,5],2), "]", sep=''))
colnames(tes3) <- colnames(coe)
```

```{r}
coe <- data.frame(apply(coe, 2, as.character), stringsAsFactors=FALSE)
ed2 <- data.frame(apply(ed2, 2, as.character), stringsAsFactors=FALSE)
tes3 <- data.frame(apply(tes3,2, as.character), stringsAsFactors=FALSE)
```

```{r}
table.2 <- data.frame(rbind(coe, ed2, tes3))

table.2
```

```{r}
table.2 <- xtable(table.2)

print(table.2, type = "html",
      file = paste("tables/", product, "_ECM.doc", sep = ""))
```

