---
title: "APT Pork"
author: "Mykola Dereva"
date: "1/24/2022"
output: html_document
---



```{r}
rm(list = ls())
```

```{r}
d <- readRDS("data/clean_data.RDS")
```


# Modeling 

```{r message=FALSE}
library(apt)
library(urca)
library(dplyr)

library(xtable)
```



There is a conflict with the dplyr lag() so I have to change it to the default 
function, so the apt package works properly 

```{r}
lag <- stats::lag
```



```{r}
product <- "Wheat flour"

prod <- x <- d %>% 
  filter(Type == "Producer",
         Product == product,
         Date >= "2013-01-01") %>% 
  select(Price) %>% 
  log() %>% 
  ts(start = c(2013,1), frequency = 12)
  
cons <- y <- d %>% 
  filter(Type == "Consumer",
         Product == product,
         Date >= "2013-01-01") %>% 
  select(Price) %>% 
  log() %>% 
  ts(start = c(2013, 1), frequency = 12)

x.name <- "Producer"
y.name <- "Consumer"

table_path <- paste("tables/", product, "/", sep = "")
```




```{r}
par(mfrow=c(2,1))
plot(x, main = x.name)
plot(y, main = y.name)
```




```{r}
ca.jo(cbind(y,x), type = "eigen", ecdet = "none") %>% summary()
```


```{r}
# Two step  cointegration test

LR <- lm(y ~ x)
# residuals
res <- ts(residuals(LR), start=start(x), end=end(x), frequency = 12)

ur.df(y = res, type = "none",  selectlags = "AIC") %>% summary()
```





```{r}
# Test for structural brakes in TS
br.x <- ur.za(x, model = "both", lag = 1)
summary(br.x)
```

```{r}
plot(br.x)
```

```{r}
br.y <- ur.za(y, model = "both", lag = 1)
summary(br.y)
```
```{r}
plot(br.y)
```





```{r}
## convenience function to make DF stationarity test tables quickly 

get_df_test_result <- function(ts.name, ts.data, 
                            test.type = c("none", "drift", "trace")) {
  
  t <- ur.df(y = ts.data, type = test.type,  selectlags = "AIC")
  
  table <- cbind(ts.name,
                  t@model,
                  t@teststat %>% t() %>% round(2),
                  t@cval,
                  t@lags) %>% as.data.frame() %>% 
  tibble::rownames_to_column()
  
  colnames(table) <- c("Parameter", "Item", "Test Type", "Test Statistics",
                     "1pct", "5pct", "10pct", "Lag")
  
  table <- table %>% select(Item, `Test Type`, Parameter, everything())
  
  return(table)
  
}
```





```{r}
LR <- lm(y ~ x)
# residuals
res <- ts(residuals(LR), start=start(x), end=end(x), frequency = 12)
```


```{r}
stationarity.table <- 
  rbind(get_df_test_result("Consumer", cons, test.type = "none"),
      get_df_test_result("Consumer", cons, test.type = "drift"),
      get_df_test_result("Consumer", cons, test.type = "trend"),
      get_df_test_result("Producer", prod, test.type = "none"),
      get_df_test_result("Producer", prod, test.type = "drift"),
      get_df_test_result("Producer", prod, test.type = "trend"),
      get_df_test_result("Coinegration", res, test.type = "none"),
      get_df_test_result("Coinegration", res, test.type = "drift"),
      get_df_test_result("Coinegration", res, test.type = "trend")) 

stationarity.table
```

```{r}
stationarity.table %>% 
    xtable() %>% 
      print(type = "html",
            file = paste(table_path,
                         product, "_Stationarity_no_diff.doc", sep = ""))
```



Both variables not stationary together, or not coinegrated 
So we need to make a first difference to stationarize them 

## First diff

```{r}
prod.diff <- y.diff <- diff(cons, differences = 1)
cons.diff <- x.diff <- diff(prod, differences = 1)
```







```{r}
stationarity.table_diff <- 
  rbind(get_df_test_result("Consumer", cons.diff, test.type = "none"),
      get_df_test_result("Consumer", cons.diff, test.type = "drift"),
      get_df_test_result("Consumer", cons.diff, test.type = "trend"),
      get_df_test_result("Producer", prod.diff, test.type = "none"),
      get_df_test_result("Producer", prod.diff, test.type = "drift"),
      get_df_test_result("Producer", prod.diff, test.type = "trend"))

stationarity.table_diff
```



```{r}
stationarity.table_diff %>% 
    xtable() %>% 
      print(type = "html",
            file = paste(table_path,
                         product, "_Stationarity_first_diff.doc",
                       sep = ""))
```






## structural brake

Test stationarity with exogenous structural break
```{r}
t <- ur.za(y = cons, model = "both",  lag = 1)
```

```{r}
summary(t)
```


```{r}

t <- ur.za(y = cons, model = "both",  lag = 1)

table <- cbind("Consumer",
                t@model,
                t@teststat %>% round(2),
                t@cval %>% t(),
                t@lag) %>% as.data.frame() 

colnames(table) <- c("Item", "Test Type", "Test Statistics",
                   "1pct", "5pct", "10pct", "Lag")

table <- table %>% select(Item, `Test Type`, everything())

table
```



```{r}
## convenience function to make Zivot and Andrews stationarity test tables quickly 

get_za_test_result <- function(ts.name, ts.data, 
                            model = c("intercept", "trend", "both"),
                            lag = 1) {
  
  t <- ur.za(y = ts.data, model = model,  lag = lag)

  table <- cbind(ts.name,
                  t@model,
                  t@teststat %>% round(2),
                  t@cval %>% t(),
                  t@lag) %>% as.data.frame() 
  
  colnames(table) <- c("Item", "Test Type", "Test Statistics",
                     "1pct", "5pct", "10pct", "Lag")
  
  table <- table %>% select(Item, `Test Type`, everything())
  
  return(table)
}
```

```{r}
get_za_test_result(ts.name = "Consumer", ts.data = cons, model = "both")
```

```{r}
stationarity.table_za <- 
  rbind(
        get_za_test_result("Consumer", cons, model = "intercept"),
        get_za_test_result("Consumer", cons, model = "trend"),
        get_za_test_result("Consumer", cons, model = "both"),
        get_za_test_result("Producer", prod, model = "intercept"),
        get_za_test_result("Producer", prod, model = "trend"),
        get_za_test_result("Producer", prod, model = "both")
        )

stationarity.table_za
```


```{r}
stationarity.table_za %>% 
    xtable() %>% 
      print(type = "html",
            file = paste(table_path,
                         product, "_Stationarity_ZA.doc",
                       sep = ""))
```





## Perron Unit Root test

```{r}
aTSA::pp.test(cons, lag.short = TRUE, type = "Z_rho")
```

```{r}
aTSA::pp.test(cons.diff, lag.short = TRUE, type = "Z_rho")
```


```{r}
aTSA::pp.test(prod, lag.short = TRUE, type = "Z_rho")
```

```{r}
aTSA::pp.test(prod.diff, lag.short = TRUE, type = "Z_rho")
```

```{r}

aTSA::pp.test(res, lag.short = F, type = "Z_rho")
```


# TAR + Cointegration


### TAR

```{r}
t3 <- ciTarThd(y=y, x=x, model="tar", lag=0)
```

```{r}
(th.tar <- t3$basic)
```


```{r}
plot(t3)
```




Choose threshold

```{r}
for (i in 1:12) { # 20 seconds
  t3a <- ciTarThd(y=y, x=x, model="tar", lag=i)
  th.tar[i+2] <- t3a$basic[,2]
}
```



```{r}
th.tar
```




### MTAR treshold

```{r}
t4 <- ciTarThd(y=y, x=x, model="mtar", lag=0)
(th.mtar <- t4$basic)
```


```{r}
plot(t4)
```



```{r}
for (i in 1:12) {
  t4a <- ciTarThd(y=y, x=x, model="mtar", lag=i)
  th.mtar[i+2] <- t4a$basic[,2]
}
th.mtar
```

The following threshold values are specific to this data. They HAVE TO be
revised for another data set. Otherwise, various errors will occur.

```{r}
(t.tar <- th.tar$V3[2])    #Best threshold for tar
(t.mtar <- th.mtar$V3[2])
```


## Lag selection 

```{r}
# max lag 
mx <- 12
```



### tar, thresold 0
```{r}
g1 <- ciTarLag(y=y, x=x, model="tar", maxlag=mx, thresh = 0)
g1
```


#### MTAR, threshlod 0
```{r}
(g2 <-ciTarLag(y=y, x=x, model="mtar",maxlag=mx, thresh=0))
```



```{r}
plot(g1)
```





```{r}
plot(g2)
```



```{r}
(g3 <-ciTarLag(y=y, x=x, model="tar", maxlag=mx, thresh=t.tar))

```



```{r}
(g4 <-ciTarLag(y=y, x=x, model="mtar",maxlag=mx, thresh=t.mtar))

```


# Table 3 Results of EG and threshold cointegration tests

Choose Lag
```{r}
vv <- 0
```



```{r}
(f1 <- ciTarFit(y=y, x=x, model="tar", lag=vv, thresh=0))
```



```{r}
(f2 <- ciTarFit(y=y, x=x, model="tar", lag=vv, thresh=t.tar ))
```



```{r}
(f3 <- ciTarFit(y=y, x=x, model="mtar", lag=vv, thresh=0 ))
```


```{r}
(f4 <- ciTarFit(y=y, x=x, model="mtar", lag=vv, thresh=t.mtar))
```




#### summary table

```{r}
r0 <- cbind(summary(f1)$dia, summary(f2)$dia,
            summary(f3)$dia, summary(f4)$dia)

diag <- r0[c(1:4, 6:7, 12:14, 8, 10, 9, 11), c(1,2,4,6,8)]

rownames(diag) <- 1:nrow(diag)
diag
```


```{r}
e1 <- summary(f1)$out
e2 <- summary(f2)$out
e3 <- summary(f3)$out
e4 <- summary(f4)$out

```


```{r}
ee <- list(e1, e2, e3, e4)
vect <- NULL


for (i in 1:4) {
  ef <- data.frame(ee[i])
  vect2 <- c(paste(ef[3, "estimate"], ef[3, "sign"], sep=""),
  paste("(", ef[3, "t.value"], ")", sep=""),
  paste(ef[4, "estimate"], ef[4, "sign"], sep=""),
  paste("(", ef[4, "t.value"], ")", sep=""))
  vect <- cbind(vect, vect2)
}

```


```{r}
item <- c("pos.coeff","pos.t.value", "neg.coeff","neg.t.value")
ve <- data.frame(cbind(item, vect))
colnames(ve) <- colnames(diag)
( res.CI <- rbind(diag, ve)[c(1:2, 14:17, 3:13), ] )
rownames(res.CI) <- 1:nrow(res.CI)

res.CI
```


### save summary table
```{r}
table <- xtable(res.CI)
print(table, type = "html",
      file = paste(table_path,
                   product, ".doc", sep = ""))
```




# APT + ECM



```{r}
(sem <- ecmSymFit(y=prod, x=cons, lag=1))  
```
```{r}
summary(sem)
```


```{r}

ccc <- summary(sem)

tr <- nrow(ccc)
ce <- tr / 2

coe <- cbind(as.character(ccc[1:ce, 2]),
    paste(ccc[1:ce, "estimate"], ccc$signif[1:ce], sep=""), ccc[1:ce, "t.value"],
    paste(ccc[(ce+1):tr,"estimate"], ccc$signif[(ce+1):tr],sep=""), ccc[(ce+1):tr,"t.value"])

colnames(coe) <- c("item",
                   paste(x.name, ".est", sep = ""),
                   paste(x.name, ".t", sep = ""),
                   paste(y.name, ".est", sep = ""),
                   paste(y.name, ".t", sep = "")
                   )

coe %>% data.frame()
```


```{r}
edia <- ecmDiag(sem, 3)
ed <- edia[c(1,6,7,8,9, 10), ]
ed2 <- cbind(ed[,1:2], "_", ed[,3], "_")
colnames(ed2) <- colnames(coe)
ed2
```



```{r}
table.1.1 <- data.frame(rbind(coe, ed2))

table.1.1
```


```{r}
table.1.1 %>%  
  xtable() %>% 
  print(type = "html",
      file = paste(table_path, product, "_Symetric_ECM.doc", sep = ""))
```



# Fitting Assymetric ECM


```{r}
aem <- ecmAsyFit(y=prod, x=cons, lag=1, model="tar",
                 split=TRUE, thresh=t.tar)
aem
```

```{r}
summary(aem)
```







```{r}
ccc <- summary(aem)
tr <- nrow(ccc)
ce <- tr / 2



coe <- cbind(as.character(ccc[1:ce, 2]),
    paste(ccc[1:ce, "estimate"], ccc$signif[1:ce], sep=""),
    ccc[1:ce, "p.value"],
    paste(ccc[(ce+1):tr,"estimate"], ccc$signif[(ce+1):tr],sep=""),
    ccc[(ce+1):tr,"p.value"])

colnames(coe) <- c("item",
                   paste(x.name, ".est", sep = ""),
                   paste(x.name, ".pval", sep = ""),
                   paste(y.name, ".est", sep = ""),
                   paste(y.name, ".pval", sep = "")
                   )

coe
```


```{r}
edia <- ecmDiag(aem, 3)
ed <- edia[c(1,6,7,8,9), ]
ed2 <- cbind(ed[,1:2], "_", ed[,3], "_")
colnames(ed2) <- colnames(coe)
ed2
```




```{r}
tes <- ecmAsyTest(aem)$out
tes2 <- tes[c(2:7,1), -1]   ### change depending on the lag number

tes3 <- cbind(as.character(tes2[,1]),
              paste(tes2[,2], tes2[,6], sep=''),
              paste("[", round(tes2[,4],2), "]", sep=''),
              paste(tes2[,3], tes2[,7], sep=''),
              paste("[", round(tes2[,5],2), "]", sep=''))

colnames(tes3) <- colnames(coe)

tes3
```

```{r}
coe <- data.frame(apply(coe, 2, as.character), stringsAsFactors=FALSE)
ed2 <- data.frame(apply(ed2, 2, as.character), stringsAsFactors=FALSE)
tes3 <- data.frame(apply(tes3,2, as.character), stringsAsFactors=FALSE)
```

```{r}
table.2 <- data.frame(rbind(coe, ed2, tes3))

table.2
```


```{r}
table.2 %>% 
  xtable() %>% 
  print(type = "html",
      file = paste(table_path, product, "_Asymetric_ECM.doc", sep = ""))
```

